---
title: "Practical Machine Learning Project - Activity Prediction"
author: "Kalil Figueroa"
date: "May 27, 2017"
output: html_document
---

## SYNOPSIS
The objective and goal of this project is to predict the manner in which 6 
participants performed several exercises using machine learning classification 
of accelerometers data on the belt, forearm, arm, and dumbell movements. The 
outcome variable in the training data  is under "classe".The "classe" variable 
classifies the correct and incorrect outcomes of A, B, C, D, and E categories.  
We will predict 20 different test cases with our model. The data for this 
project come from: http://groupware.les.inf.puc-rio.br/har.

## Libraries
```{r}
library(caret)
library(randomForest)

set.seed(2017)
```

## Reading Data
```{r}
a <- read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!", "") )
test <- read.csv("pml-testing.csv")
dim(a)
dim(test)
```

## Cleaning Data

### Take out vars with more than 95% of na values
```{r}
a <- a[, colMeans(is.na(a)) < 0.05]
dim(a)
```

### Take out vars with near zero variance
```{r}
nzv <- nearZeroVar(a, saveMetrics = FALSE)
a <- a[ , -nzv]
dim(a)
```

### Take out ID and timestamp columns
```{r}
colnames(a[1:5])
a <- a[,-c(1:5)]
dim(a)
```

## Data Partition 70/30 for training and validation
```{r}
partition <- createDataPartition(a$classe, p = 0.7, list = FALSE)
trainData<- a[partition,]
validData <- a[-partition,]
```

## Train the model - 10-fold cross validation - random forest
We will use the random forest model because it is widely used and has high accuracy.
```{r}
fit <- train(classe ~ ., data = trainData, ntree = 100, method = 'rf',
                 trControl = trainControl(method = "cv", number = 10))
fit$finalModel
```


```{r}
plot(fit$finalModel)
```

With an out of sample error rate of 0.3% we can move confidently to the 
validation set.

## High accuracy - check model with validation set
```{r}
predictFit <- predict(fit, validData)
confusionMatrix(predictFit, validData$classe)
```

Out of sample error is 1 - 0.9973 = 0.0027. We are confident with our final model.

## High accuracy - move on to test set prediction
```{r}
predictTest <- predict(fit, test)
print(predictTest)
```
